---
title: "Accommodation"
author: "sara"
format: 
  html:
    embed-resources: true
---

```{r}
library(dplyr)
library(ggplot2)
library(here)
library(scales)

purrr::map(list.files(here("R"), full.names = T), source)
```



# Context

Travelling in Iceland, I want to learn something from a numbers perspective as well. here, I am examining the data available for accommodation.

The Statistical Office of Icceland has a database that is accessible via an [API](https://www.scb.se/api_en/) that seems to come from the Swedish Statistical office, and handily enough has an [R package](https://github.com/rOpenGov/pxweb) as well :-) 

# Total number of overnight stays in Iceland over time

```{r}
df_region <- get_accommodation_by_region()

df_aggr <- df_region |> 
  group_by(year, unity) |> 
  summarise(value = sum(value))
```

```{r}
df_aggr |> 
  ggplot(aes(x = year, y = value, color = unity)) +
  geom_line() +
  scale_y_continuous(labels = label_number()) +
  labs(title = "Number of arrivals and overnight stays in Iceland",
       color = "",
       y = "")
```


# by month

```{r}
df_aggr_month <- df_region |> 
  group_by(year, unity, month) |> 
  summarise(value = sum(value))
```

```{r}
df_aggr_month |> 
  filter(unity == "Overnight stays") |> 
  ggplot(aes(x = month, y = value, color = year, group = year)) +
  geom_line() +
  scale_y_continuous(labels = label_number()) +
  labs(title = "Number of overnight stays in Iceland by month",
       color = "",
       y = "")
```


# by region

# by origin
